Assessment and optimization of respiratory syncytial virus prophylaxis regimens in Connecticut, \Sexpr{formatYear(startDate)}-\Sexpr{formatYear(endDate)}

### 2019-03-01 Paper/RSVOptimization.rnw
Supplement improvements

<<util, warning=FALSE>>=
@

@@

% TODO figure showing prophylaxis regimens on a calendar?

@@

@@

A simple example of this is our estimation of RSV season onset. We begin the process by defining season onset to occur when cumulative incidence rises above \Sexpr{seasonThreshold * 100}\% of total cases, and obtaining a $P$-spline approximation to our data (using the \texttt{mgcv} R package). This approximation is characterized by a matrix of coefficients, each of which is normally distributed with a known mean and variance.

We randomly sample each coefficient from its distribution, thereby obtaining a specific spline that smoothly approximates our data. We calculate RSV season onset for this specific spline, and then repeat the coefficient sampling process, thereby iteratively sampling RSV season onset from its (unknown) distribution.

@@

More interestingly, the same approach can be used to estimate time-series outcomes. We use this in our computation of preventable cases of RSV. We begin by defining a preventable case of RSV to be a case of RSV that occurs while prophylaxis (under a specific set of prophylaxis guidelines) offers protection; that is, a case that would have been prevented by administration of prophylaxis under those guidelines, if the prophylaxis were 100\% effective. As before, we obtain a $P$-spline approximation to our data.

We then sample $P$-spline coefficients from their distributions to obtain a specific approximation to our data, and compute the corresponding time series of preventable RSV cases. Repeating the coefficient-sampling process yields time series of preventable RSV cases sampled from their (again unknown) distribution.

This process can also be generalized to any time-series outcome of an infectious disease outbreak or season.

Furthermore, the two approaches can be combined to calculate complex non-linear outcomes. We use this to estimate the fraction of preventable RSV cases in an RSV season, defined as the number of preventable cases divided by the total number of cases. The preventable fraction is a non-linear function of the RSV time series; as a result, interval estimation of the preventable fraction cannot be obtained by point-wise estimation of the RSV time, followed by applying a function to point-wise estimates. Instead, estimates of the entire time series must be sampled, and the preventable fraction then calculated for each sampled time series.

Here we follow a similar process, wherein we sample coefficients from their distributions, use the resultant $P$-spline to calculate the all-case time series as well as the preventable-case time series, from which we then calculate the preventable fraction; the end result is that we sample the preventable fraction from its (unknown) distribution.

>>>>

<<paper, warning=FALSE>>=
@

@@

@@

<<supplement_libraries, cache=FALSE>>=
@

<<supplement, cache=TRUE>>=
@

@@

A simple application of this is our estimation of RSV season onset. Consider, for example, RSV-associated hospitalizations in Tolland County, Connecticut in Figure \ref{fig:tollandCases}. (This is a low-incidence county, which makes it a poor candidate for our main analysis, but its large confidence intervals make it easier to visualize our methods.)

<<tollandCases, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/tollandCases.tex", figuresDir), width=pagePlotWidth * 0.4, height=pagePlotHeight, pointsize=10)

    ggplot(tollandObs) +
        theme_light(base_size=plotTextBaseSize) +
        geom_point(aes(x=time, y=rsv.cum.frac), size=.5) +
        scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=range(c(monthBoundaries$min, monthBoundaries$max)), expand=c(0, 0)) + 
        theme(
            panel.grid.major.x=element_blank(),
            legend.title=element_blank(),
            axis.title.y=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
            legend.position="bottom"
        )

\begin{figure}
\begin{center}
\input{\figuresDir/tollandCases.tex}
\caption{Cumulative relative incidence of RSV-associated hospitalizations in Tolland county, Connecticut, aggregated across years \Sexpr{formatYear(startDate)}-\Sexpr{formatYear(endDate)}}
\label{fig:tollandCases}
\end{center}
\end{figure}

<<tollandPredMini, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/tollandPredMini.tex", figuresDir), width=pagePlotWidth * 0.4, height=pagePlotHeight, pointsize=10)

    ggplot(tollandPredMini) +
      theme_light(base_size=plotTextBaseSize) +
      geom_line(aes(x=time, y=rsv.cum.frac, group=sim), color="grey") +
      geom_segment(aes(x=-Inf, y=seasonThreshold, xend=+Inf, yend=seasonThreshold), linetype="11", data=data.frame(), size=.375) + 
      geom_point(data=tollandObs, aes(x=time, y=rsv.cum.frac), size=0.5) +
      scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=range(c(monthBoundaries$min, monthBoundaries$max)), expand=c(0, 0)) + 
      theme(
        panel.grid.major.x=element_blank(),
        legend.title=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
        legend.position="bottom"
      )

<<tollandOnsetMini, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/tollandOnsetMini.tex", figuresDir), width=pagePlotWidth * 0.4, height=pagePlotHeight, pointsize=10)

    ggplot(tollandPredMini) +
      theme_light(base_size=plotTextBaseSize) +
      geom_line(aes(x=time, y=rsv.cum.frac, group=sim), color="grey") +
      geom_segment(aes(x=-Inf, y=seasonThreshold, xend=+Inf, yend=seasonThreshold), linetype="11", data=data.frame(), size=.375) + 
      geom_point(data=tollandObs, aes(x=time, y=rsv.cum.frac), size=0.5) +
      geom_point(data=tollandOnsetMini, aes(x=onset, y=seasonThreshold), size=0.75, shape=17) + 
      geom_segment(data=tollandOnsetMini, aes(x=onset, xend=onset, y=seasonThreshold, yend=0), linetype="11") + 
      scale_y_continuous(limits=c(0, 2*seasonThreshold)) + 
      scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=c(zoomedStartWeek, zoomedEndWeek), expand=c(0, 0)) + 
      theme(
        panel.grid.major.x=element_blank(),
        legend.title=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
        legend.position="bottom"
      )

\begin{figure}
\begin{center}
\begin{subfigure}{.4\textwidth}
    \begin{center}
    \input{\figuresDir/tollandPredMini}
    \caption{Five randomly sampled $P$-splines approximating the RSV time series.}
    \label{fig:tollandPredMini}
    \end{center}
\end{subfigure}
\begin{subfigure}{.4\textwidth}
    \begin{center}
    \input{\figuresDir/tollandOnsetMini}
    \caption{Season onsets derived from five randomly sampled approximating splines.}
    \label{fig:tollandOnsetMini}
    \end{center}
\end{subfigure}
\end{center}
\caption{Cumulative relative incidence of RSV-associated hospitalizations in Tolland county, Connecticut, with five randomly sampled approximating $P$-splines.}
\end{figure}

We begin the process by defining season onset to occur when cumulative incidence rises above \Sexpr{seasonThreshold * 100}\% of total cases, and obtaining a $P$-spline approximation to our data (using the \texttt{mgcv} R package). This approximation is characterized by a matrix of coefficients, each of which is normally distributed with a known mean and variance.

We randomly sample each coefficient from its distribution, thereby obtaining a specific spline that smoothly approximates our data. Five such randomly sampled splines are shown in Figure \ref{fig:tollandPredMini}. We then calculate RSV season onset for each sampled spline (Figure \ref{fig:tollandOnsetMini}).

Repeating this process many times (\Sexpr{simulations}, in our main analysis) then allows us to estimate the distribution of RSV season onset and therefore make interval estimates. In the case of Tolland county, the result of such season onset sampling is shown in Figure \ref{fig:tollandOnsetFull}.

<<tollandOnsetFull, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/tollandOnsetFull", figuresDir), width=pagePlotWidth * 0.4, height=pagePlotHeight, pointsize=10)

    ggplot(tollandPredFull) +
      theme_light(base_size=plotTextBaseSize) +
      geom_ribbon(aes(x=time, ymin=min, ymax=max), fill="gray", na.rm=FALSE) +
      geom_segment(aes(x=-Inf, y=seasonThreshold, xend=+Inf, yend=seasonThreshold), linetype="11", data=data.frame(), size=.375) + 
      geom_point(data=tollandObs, aes(x=time, y=rsv.cum.frac), size=0.5) +
      geom_violinh(
        data=tollandOnsetFull,
        aes(x=onset, y=seasonThreshold),
        width=.01,
        fill="gray50", color="black",
        trim=FALSE, draw_quantiles=c(0.025, 0.5, 0.975)
      ) +
      scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, expand=c(0, 0)) +
      scale_y_continuous() +
      coord_cartesian(xlim=c(zoomedStartWeekFull, zoomedEndWeekFull), ylim=c(0, 4*seasonThreshold)) +
      theme(
        panel.grid.major.x=element_blank(),
        legend.title=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
        legend.position="bottom"
      )
    dev.off()
@

\begin{figure}
\begin{center}
\input{\figuresDir/tollandOnsetFull}
\caption{Distribution of approximating $P$-splines and the calculated season onset with $N=\Sexpr{simulations}$ samples, for cumulative relative incidence of RSV-associated hospitalizations in Tolland county, Connecticut.}
\label{fig:tollandOnsetFull}
\end{center}
\end{figure}

@@

Furthermore, because our approach involves sampling approximations to the observed time series as a whole, rather than estimating the time series at isolated points in time, we are able to produce interval estimates of arbitrarily complex outcome measures.

We take advantage of this in our estimation of the preventable fraction of RSV cases. Here we begin by defining a preventable case of RSV to be a case of RSV that occurs while prophylaxis offers protection, and we define the preventable fraction of a prophylaxis administration strategy to be the ratio of preventable cases under that prophylaxis strategy to the total cases of RSV.

As before, we then sample $P$-spline coefficients to obtain a specific $P$-spline approximation, and compute preventable fraction for each sampled $P$-spline in order to obtain the distribution of preventable fractions.
